Towards practical dynamic induction control of wind farms: analysis of optimally controlled wind-farm boundary layers and sinusoidal induction control of first-row turbines

### 2018-02-14 draft.tex
Added all Johans comments, version WM2_JM1

In the remainder of this section, time-averaged wind-farm flow properties will be investigated. Here and throughout the remainder of this paper, we focus on case C3t5, as it produces similar energy gains as the highest-yield case C3t0 (see Figure \ref{fig:power_MM17}a), but achieves this with smooth thrust coefficient signals. In the following discussion, the time averaging operation is denoted by an overline, and flow field variables are decomposed into mean and fluctuating components as $\widetilde{u} = \overline{\widetilde{u}} + \widetilde{u}' \equiv \widetilde{U} + \widetilde{u}'  $. Figure \ref{fig:flowfield_statistics} illustrates time-averaged flow field quantities of the reference case (left panels, a1 -- g1) and the differences between the optimized C3t5 case and the reference case (right panels, a2 -- g2). Simulation results are averaged over the different columns and are shown as either topviews at hub height (Figures \ref{fig:flowfield_statistics}b,f,g) or sideviews through a turbine column (Figures \ref{fig:flowfield_statistics}a,c,d,e).

@@

The observations and experiments from previous sections have revealed information that increases the understanding the optimized thrust coefficients, and can be used as a starting point towards designing practical wind-farm controllers that do not require computationally prohibitive LES-based optimal control simulations. A first conclusion that can be drawn from the analysis above is that wind turbines can be classified into three distinct categories, based on their position within the farm: first-row turbines, last-row turbines, and intermediate turbines.

The most salient behavior can be found in the first-row turbines (R1). Firstly, Figure~\ref{fig:spectra_thrust_cthat} showed that these turbines exhibit the largest variability in thrust coefficients and resulting thrust forces. Second, Figure~\ref{fig:multirow}a has shown that the optimized thrust coefficients in R1 are the only ones that can be applied independently from the other rows, and Figure~\ref{fig:single_row_opt} illustrated that the first row holds the greatest potential for power optimization of all rows in the farm. Third, the first row is the only one that has been shown to retain part of its power increase after eliminating the correlation between the optimized thrust coefficients and the flow field in Figure~\ref{fig:scrambled}, suggesting a flow-invariant component in the optimized coefficients. Finally, it is worth remarking the observed peak around $fD/U_\infty \approx 0.2 \dots 0.3$ from Figure~\ref{fig:spectra_thrust_cthat}. It will be shown below that variations in the thrust coefficient are related to specific flow events in the wake of R1.

The characteristics of the last row turbines (R12) also stand out from the rest. This is mainly caused by the fact that, by definition, the last row has no downstream turbines for which it can generate more favorable flow conditions, causing them to operate greedily in an optimal control setting. This explains the relatively quiescent behavior of optimized thrust coefficients in Figure~\ref{fig:controls}, and the limited potential for power increase in Figure~\ref{fig:single_row_opt}.

The remaining intermediate rows (R2--R11) have similar spectral thrust characteristics (Figure~\ref{fig:spectra_thrust_cthat}) and potential for power increase (Figure~\ref{fig:single_row_opt}), situated in between but clearly separated from first- and last-row turbines. Further, it is worth noting that the behavior and analysis of control actions in these turbines is most complex: not only do they influence downstream turbines, they in turn are dependent on controls of upstream turbines.

A second conclusion is that, whether or not the wind-farm is controlled with the possibility of active response and cooperation between turbines, the resulting power and thrust characteristics are very similar. Firstly, both Figure~\ref{fig:multirow}b (full-farm optimization) and Figure~\ref{fig:gains_singlerow} (single-row optimization) suggest that self-optimization of turbines is negligible, and that power gains are mainly achieved by generating more favorable flow conditions for downstream turbines. Furthermore, the remarkable match between optimized thrust coefficients for full-farm and single-row optimal control cases from Figure~\ref{fig:control_C3t5_singlerow} indicates that, at least for the first row, upstream controls do not rely on active downstream response. Instead, Figure~\ref{fig:multirow}b shows that, for any row $i$, the full potential in power increase is virtually attained by applying controls only for the upstream turbines up until row $i-1$. Furthermore, Figure~\ref{fig:multirow}a illustrates that optimized thrust coefficients in downstream rows by themselves do not result in power increase without also applying optimized coefficients in their upstream neighbors. This indicates that controls are tuned to specific flow events. This also explains the discrepancies between full-farm and single-row optimized thrust coefficients for downstream rows in Figure~\ref{fig:control_C3t5_singlerow}. 
These observations strongly suggest that the optimized thrust coefficients are designed in a parabolic manner, i.e. with a unidirectional propagation of control information from the first row to the last, and very little upstream influence of downstream turbine actions.

With this in mind, the following section of this paper will focus on the first and most promising link of the control chain: the turbines situated in the first row of the wind farm.

\section{First-row turbine behavior}\label{sec:analysis_flow_vis}
As shown in the previous section, the behavior of turbines within the first upstream row stands out from the others. Firstly, they are not influenced by any dynamic actions of the upstream turbines which makes their behavior somewhat easier to study. Secondly, they are unwaked and hence subjected to completely different flow characteristics than their downstream counterparts leading to much larger power extraction in the first row than downstream. Third, it was shown that the first-row controls show some inherent characteristics that are not tuned to specific flow events, and hence hold promise for simple and universally applicable control strategies. Therefore, the current section further focuses on the analysis of the first row.

First, a qualitative analysis of the instantaneous flow field is performed in Section~\ref{sec:opt_visualization}, resulting in the observation of vortex rings being shed from first-row turbines. Thereafter, this mechanism is mimicked by imposing sinusoidally varying thrust coefficients in these turbines in Section~\ref{sec:opt_sinus}, with the aim of increasing power through similar mechanisms as in the computationally expensive optimal control cases.

@@

Figure~\ref{fig:sinus_fullscale} shows the power extraction of the full-scale test. Figure~\ref{fig:sinus_fullscale}a and~\ref{fig:sinus_fullscale}b show the relative power gains over the reference case for the full wind farm and the first four turbine rows respectively. A first observation is that the optimal pair $(St^\star, A^\star)$ differs for both measures, suggesting that the increased power in front rows leads to losses in the downstream rows. This is confirmed by the row-wise power shown in Figure~\ref{fig:sinus_fullscale}c: from the fifth row onwards, power is slightly reduced in the sinusoidal cases. This explains the negligible increase in total farm power observed in Figure~\ref{fig:sinus_fullscale}a.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{gains_fullscale.eps}
	\caption{Power extraction of full-scale sinusoidal thrust case ($S = 6D$, $z_0 = 10^{-1}$ m). \emph{a) } Relative gain in mean wind-farm power extraction over reference case as a function of sine amplitude $A$ and frequency $St$. \emph{b) } Relative gain in mean power extraction for the first four rows. \emph{c) } Row-averaged mean power extraction for the best sinusoidal case, the optimal control case C3t5 from MM17, and the reference case, normalized by first-row reference power.\label{fig:sinus_fullscale} }
\end{figure}

>>>>

In the remainder of this section, time-averaged wind-farm flow properties will be investigated. Here and throughout the remainder of this paper, we focus on case C3t5, as it produces similar energy gains as the highest-yield case C3t0 (see Figure \ref{fig:power_MM17}a), but achieves this with smooth thrust coefficient signals. In the following discussion, the time averaging operation is denoted by an overline, and flow field variables are decomposed into mean and fluctuating components as $\widetilde{u} = \overline{\widetilde{u}} + \widetilde{u}' \equiv \widetilde{U} + \widetilde{u}'  $. Figure~\ref{fig:flowfield_statistics} illustrates time-averaged flow field quantities of the reference case (left panels, a1 -- g1) and the differences between the optimized C3t5 case and the reference case (right panels, a2 -- g2). Simulation results are averaged over the different columns and are shown as either topviews at hub height (Figures \ref{fig:flowfield_statistics}b,f,g) or sideviews through a turbine column (Figures \ref{fig:flowfield_statistics}a,c,d,e).

@@

The observations and experiments from previous sections have revealed information that increases the understanding the optimized thrust coefficients, and can be used as a starting point towards designing practical wind-farm controllers that do not require computationally expensive LES-based optimal control simulations.

A first conclusion is that wind turbines can be classified into three distinct categories, based on their position within the farm: first-row turbines, last-row turbines, and intermediate turbines. The most salient behavior can be found in the first-row turbines (R1). It was shown that these turbines exhibit the largest variability in thrust forces and hold the greatest potential for power optimization. Furthermore, they are not influenced by upstream turbine control action, and are the only turbines that retain part of the power gains after eliminating possible correlation between controls and specific flow events.  
The characteristics of the last row turbines (R12) also stand out from the rest due to the fact that, by definition, the last row has no downstream turbines and hence holds no further potential for coordinated control. 
The remaining intermediate rows (R2--R11) have similar spectral thrust characteristics and potential for power increase, situated in between but clearly separated from first- and last-row turbines. Further, it is worth noting that the behavior and analysis of control actions in these turbines is most complex: not only do they influence downstream turbines, they in turn are dependent on controls of upstream turbines.

A second conclusion is that, whether or not the wind-farm is controlled with the possibility of active response and cooperation between turbines, the resulting power and thrust characteristics are very similar. It was shown that self-optimization is very limited, and that, for any row $i$, the full potential in power increase is virtually attained by applying controls only for the upstream turbines up until row $i-1$. These observations strongly suggest that the optimized thrust coefficients are designed in a parabolic manner, i.e. with a unidirectional propagation of control information from the first row to the last, and very little upstream influence of downstream turbine actions.
With this in mind, the following section of this paper will focus on the first and most promising link of the control chain: the turbines situated in the first row of the wind farm.

\section{First-row turbine behavior}\label{sec:analysis_flow_vis}
%As shown in the previous section, the behavior of turbines within the first upstream row stands out from the others. Firstly, they are not influenced by any dynamic actions of the upstream turbines which makes their behavior somewhat easier to study. Secondly, they are unwaked and hence subjected to completely different flow characteristics than their downstream counterparts leading to much larger power extraction in the first row than downstream. Third, it was shown that the first-row controls show some inherent characteristics that are not tuned to specific flow events, and hence hold promise for simple and universally applicable control strategies. Therefore, the current section further focuses on the analysis of the first row.

The current section further focuses on the analysis of the first row turbines. First, a qualitative analysis of the instantaneous flow field is performed in Section~\ref{sec:opt_visualization}, resulting in the observation of vortex rings being shed from first-row turbines. Thereafter, this mechanism is mimicked by imposing sinusoidally varying thrust coefficients in these turbines in Section~\ref{sec:opt_sinus}, with the aim of increasing power through similar mechanisms as in the computationally expensive optimal control cases.

@@

Figure~\ref{fig:sinus_fullscale} shows the power extraction of the full-scale test. Figure~\ref{fig:sinus_fullscale}a shows the relative power gains over the reference case for the full wind farm. It can be seen that the total power gain or loss is below 0.5\% for each of the sinusoidal control cases. Figure~\ref{fig:sinus_fullscale}b shows the row-wise power extraction for the reference case, the sinusoidal thrust case with $(St, A) = (0.25, 1.5)$, and the optimal control case C3t5. It is shown that, although the second row of the sinusoidal thrust case achieves similar power gains as those observed above, from the fifth row onwards power is slightly reduced in the sinusoidal case.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{gains_fullscale.eps}
	\caption{Power extraction of full-scale sinusoidal thrust case ($S = 6D$, $z_0 = 10^{-1}$ m). \emph{a) } Relative gain in mean wind-farm power extraction over reference case as a function of sine amplitude $A$ and frequency $St$. \emph{b) } Row-averaged mean power extraction for the best sinusoidal case, the optimal control case C3t5 from MM17, and the reference case, normalized by first-row reference power.\label{fig:sinus_fullscale} }
\end{figure}
